### 2013-10-01 Publications/2013-ISBI14/01-abstract.tex
Final version

\begin{abstract}
Connectivity analysis on diffusion MRI data of the whole-brain
suffers from distortions caused by the standard 
echo-planar imaging acquisition strategies. %
%These artifacts are present at interfaces between tissues,
%due to the inhomogeneity of the field derived from the 
% discontinuity of magnetic susceptibility.
These images show characteristic geometrical deformations
and signal destruction that are a significant pitfall
limiting the success of tractography algorithms.

Several retrospective correction techniques are readily
available. In this work, we use a digital phantom designed
for the evaluation of connectivity pipelines. We subject
the phantom to a ``theoretically correct'' and plausible
deformation that resembles the artifact under investigation.
We correct data back, with three standard methodologies
(namely fieldmap-based, reversed encoding-based, and
registration-based). Finally, we rank the methods based
on their geometrical accuracy, the dropout compensation,
and their impact on the resulting tractography.
\end{abstract}%
\begin{keywords}
susceptibility artifacts, diffusion MRI, tractography, connectivity.
\end{keywords}

>>>>

\begin{abstract}
Connectivity analysis on diffusion MRI data of the whole-brain
suffers from distortions caused by the standard 
echo-planar imaging acquisition strategies. %
%These artifacts are present at interfaces between tissues,
%due to the inhomogeneity of the field derived from the 
% discontinuity of magnetic susceptibility.
These images show characteristic geometrical deformations
and signal destruction that are an important pitfall
limiting the success of tractography algorithms.

Several retrospective correction techniques are readily
available. In this work, we use a digital phantom designed
for the evaluation of connectivity pipelines. We subject
the phantom to a ``theoretically correct'' and plausible
deformation that resembles the artifact under investigation.
We correct data back, with three standard methodologies
(namely fieldmap-based, reversed encoding-based, and
registration-based). Finally, we rank the methods based
on their geometrical accuracy, the dropout compensation,
and their impact on the resulting connectivity matrices.
\end{abstract}%
\begin{keywords}
susceptibility artifacts, diffusion MRI, tractography, connectivity.
\end{keywords}

### 2013-10-01 Publications/2013-ISBI14/02-introduction.tex
Final version

Among all the difficulties that such a complex workflow
raises \cite{jones_twenty-five_2010}, we will
address here the susceptibility-derived artifacts,
for which \Gls*{epi} schemes are especially sensitive.
Magnetic susceptibility disturbs the magnetic field
close to tissue interfaces. This inhomogeneity
of the field translates in a highly distorted
anatomy and a significant signal destruction in
certain regions of the brain 
(e.g. the orbitofrontal lobe, for the proximity of the
air surrounding sinuses). This artifact has been
well described, generally within the context of
functional MRI which also uses \gls*{epi}.

Various approaches have been proposed to correct
for the disortions. \Gls*{fmb} methods
\cite{jezzard_correction_1995} rely on one extra
acquisition (\emph{field mapping}),
that probes the inhomogeneity of the B0 field.
A second theory-based breed of methodologies acquire a 
map of the \acrlong*{psf} of the \gls*{epi} readouts to correct 
for the artifact \cite{robson_measurement_1997}.
Another approach referred to as \gls*{reb}, acquires an
extra \gls*{epi} volume in the orthogonal or reversed
phase encoding that can be combined to remove the
geometric distortions
\cite{cordes_geometric_2000,chiou_simple_2000}.
The last family of methodologies acquires an
extra T2-weighted image, and uses it as an anatomical
reference to find the deformation map through nonlinear 
registration \cite{kybic_unwarping_2000,studholme_accurate_2000}.
These \gls*{t2b} methods usually map
the T2 image to the \emph{baseline} volume or \textit{b0}
of \gls*{dmri}. The choice of T2 is due to the strong
similarity of intensity distribution with respect the \textit{b0}.
More recent works report extensions or combined
approaches of existing techniques
\cite{andersson_how_2003,zaitsev_point_2004,%
holland_efficient_2010,andersson_comprehensive_2012}.

>>>>

Among all the difficulties that such a complex workflow
raises \cite{jones_twenty-five_2010}, we will
address here the susceptibility-derived artifacts,
for which \Gls*{epi} schemes are especially sensitive.
Magnetic susceptibility disturbs the magnetic field
close to tissue interfaces. This inhomogeneity
of the field translates in a highly distorted
anatomy and a significant signal destruction in
certain regions of the brain 
(e.g. the orbitofrontal lobe, for the proximity of the
air surrounding sinuses). This artifact has been
well described, generally within the context of
functional MRI, which also uses \gls*{epi}.

Various approaches have been proposed to correct
for the distortions. \Gls*{fmb} methods
\cite{jezzard_correction_1995} rely on one extra
acquisition (\emph{field mapping}),
that probes the inhomogeneity of the B0 field.
A second theory-based breed of methodologies acquire a 
map of the \acrlong*{psf} of the \gls*{epi} readouts to correct 
for the artifact \cite{robson_measurement_1997}.
Another approach referred to as \gls*{reb}, acquires an
extra \gls*{epi} volume in the orthogonal or reversed
phase encoding that can be combined to remove the
geometric distortions
\cite{cordes_geometric_2000,chiou_simple_2000}.
The last family of methodologies acquires an
extra T2-weighted image, and uses it as an anatomical
reference to find the deformation map through nonlinear 
registration \cite{kybic_unwarping_2000,studholme_accurate_2000}.
These \gls*{t2b} methods usually map
the T2 image to the \emph{baseline} volume or \textit{b0}
of \gls*{dmri}. The choice of T2 is due to the strong
similarity of intensity distribution with respect the \textit{b0}.
More recent works report extensions or combined
approaches of existing techniques
\cite{andersson_how_2003,zaitsev_point_2004,%
holland_efficient_2010,andersson_comprehensive_2012}.
